### 2016-11-17 README.Rmd
Updated README

---
title: "Readme for Muir and Angert"
author: "Chris Muir and Amy Angert"
output: html_document
---

This repository contains the source for the paper Muir CD and Angert AL. Grow with the flow: a latitudinal cline in physiology is associated with more variable precipitation in *Mimulus cardinalis*.

See the Sweave file 'ms/ms.Rnw' for code chunks. If you have R packages installed, you should be able to recreate the manuscript locally.

A preprint of this paper is available on biorxiv: http://dx.doi.org/10.1101/080952