### 2017-10-13 README.Rmd

# card-cline
Grow with the flow: a latitudinal cline in physiology is associated with more variable precipitation in *Erythranthe cardinalis*. In Press at *Journal of Evolutionary Biology*.

This project was developed by [Chris Muir](www.chrisdmuir.com) and [Amy Angert](http://angert.botany.ubc.ca/) .

More information about the study is available in a preprint which you can find on [biorxiv](https://doi.org/10.1101/080952) or on [github](https://github.com/cdmuir/card-cline/blob/master/ms/ms.pdf).

## Downloading repository

1. Download or clone this repository to your machine.
2. Open `card-cline.Rproj` in [RStudio](https://www.rstudio.com/)
3. Install R packages if necessary.

## Generating manuscript

Simply open `ms/ms.Rnw` and compile using RStudio.

[![CC BY](http://i.creativecommons.org/l/by/3.0/88x31.png)](http://creativecommons.org/licenses/by/3.0/)